Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-05-31 WRITEUP/ms.tex
revised abstract

\begin{abstract}
We study the properties of molecular cloud complexes (MCCs) using cosmological zoom-in simulation of a
$z\simeq 6$ prototypical galaxy (``\flower'').
%
We identify MCCs using an H$_2$ density-based clump finder, and compare their mass, size, velocity dispersion, gas surface density, and virial parameter
to \obs.
%
In \flower, the typical MCC mass is $10^{6.5}$\,\Msun, with sizes $R\lesssim100$\,pc, comparable to nearby spirals and starburst galaxies.
%
We identify gravitationally unstable MCCs both via Toomre $Q_{\rm eff}$ analysis and
the more observationally accessible virial parameter ($\alpha_{\rm vir}$) analysis.
MCCs in the main disk of \flower are globally stable, but their internal substructures have low $\alpha_{\rm vir}$ and $Q_{\rm eff}$ values, indicative of
instability to % gravitational
collapse.
%
Their velocity dispersion and gas surface density are systematically higher than those observed in nearby galaxies, but comparable to starburst galaxies.
%
We show that large-scale rotation is an important contributor to the velocity dispersion of some MCCs.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA) on
$\lesssim$40\,pc are necessary to shed light on the star-forming gas dynamics since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

%mm rearranged and combined this and the following paragraph to proceed from observation to conclusion
Early galaxies have higher star formation rates \citep[SFR; ][]{Behroozi13b, Sparre15a, Maiolino15a, Dunlop17a} and smaller sizes \citep[e.g.,][]{Bouwens11a, Ono13a} compared to the local population \citep[see also a review by][]{Stark16a}.
%
As a consequence, we expect them to be significantly more ionized, and have intense and hard interstellar radiation fields (ISRF). Since their metallicity and dust content are also expected to be lower in these early evolutionary stages, shielding of UV photons---responsible for the photoheating of the gas---  is strongly reduced. Such differences in turn affect the regulation of the thermal and chemical state of the multi-phase interstellar medium (ISM). Studying ISM properties of early galaxies is essential for understanding how \SF proceeds under more extreme conditions.
%
Even in the local Universe, where detailed \obs can be made, variations in molecular cloud
%       \MM{do Hughes et al. study cloud complexes, as you do, or can you derive info about complexes from their data?}
%     \DL{This is just a statement to say there are variations among local GMCs. Together with the different galaxy properties found for early galaxies, it is interesting to study properties of molecular structures at earlier epochs. }
properties have been observed between different galaxy populations (see e.g., \citealt{Hughes10a, Hughes13b}).  Given that \highz galaxies statistically represent the early evolutionary stages of present-day galaxies, it is thus reasonable to pose the question: 
{\it what are the physical properties of
molecular cloud structures in early galaxies, and how do they differ from those found in local \galpop?}
%

@@

Motivated by Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, we assume equilibrium (i.e.,\ ${\ddot{\mathcal{I}}}=0$), define the cloud surface density as
$\Sigma$\eq$M/\pi R^2$, and rewrite the previous equation as
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \frac{3}{5} \pi G \Sigma \right)\,,
\label{eqn:v0}
\end{equation}
which further reduces to
\begin{equation}
\frac{\sigma^2}{R} = \frac{\pi}{5} G \Sigma\,,
\label{eqn:SVE}
\end{equation}
if the external pressure $P_{\rm ext}=0$. For this case (often referred to as simple virial equilibrium) from the balance between kinetic and gravity terms we can define the virial parameter as
\begin{equation}
\alpha_{\rm vir} \equiv  \frac{5\sigma^2R}{GM_{\rm gas}} = \frac{5\sigma_{\rm gas}^2}{\pi G \Sigma R}\,,
\label{eqn:alpha}
\end{equation}

@@

In the clump-finding process (in position-position-position space; PPP), we employ a set of different density thresholds defined based on the molecular hydrogen density of \flower at different evolutionary stages ($z=6.0$--7.2).
%
We note that this process is the three-dimensional analog to identifying molecular structures based on the noise levels in position-position-velocity (PPV) maps that observers obtain with telescopes, using molecular line tracers such as CO, CS, and HCN. This is commonly done by
identifying clumps based on/after applying S/N-clipping, using tools such as \ncode{aips}'s task \ncode{serch}, \ncode{clumpfind}, and \ncode{cprops}; e.g., \citealt{Williams94a, Oka01a, Rosolowsky06a, Rosolowsky08a, DonovanMyer13a}).
%
%We note that, owing to the nature of \obs, such structures are identified in position-position-velocity (PPV) space, whereas in simulations, one has the full 6D spatial-kinematic information, and can therefore cleanly identify structures directly using the density field in position-position-position (PPP) space.
Existing studies find a good correspondence in the dynamical properties extracted in PPV- versus PPP-space for well-isolated structures (\citealt{Ballesteros-Paredes02a, Heitsch09a, Shetty10a, Beaumont13a, Pan15a}, but see \citealt{Ballesteros-Paredes02a} and \citealt{Shetty10a} for a discussion on caveats and limitations).
% large scale structure in PPP may be identified as numerous lower mass structures in the PPV cube due to gradients in the LOS and conversly, superposition when spatially distinct components with the same velocity are "merged" into one component in the PPV space when observed along the LOS (see Fig. 1 of Beaumount13a). (may be worth putting more thoughts on this after the report is due).

@@

\begin{figure*}
\centering
\includegraphics[trim=0 0 0 0, clip, width=\textwidth]{\figpath/3by2_clumpProp_ss16-ss22.png}
\caption{
Linewidth-size relation (top), $\alpha_{\rm vir}$-mass relation
(middle), and $\sigma^2/R$-$\Sigma_{\rm gas}$ relation (bottom) for
MCCs (star symbols) identified in the two most extreme evolutionary
stages of \flower\ --- accreting phase (left) and starburst phase
(right). Star symbols are color-coded by the density thresholds
$n_{\rm cut}$, as illustrated by the colorbar.
Literature data in the linewidth-size plots are from \citet{Heyer04a, Rosolowsky05a, Bolatto08a,
% black dot = extragalactic GMCs in quiescent environments
Leroy15a}, and \citet{Swinbank11a}, and the empirical scaling relations are
from \citet{Larson81a, Heyer04a, Bolatto08a}.
Data points in the $\alpha_{\rm vir}$-mass figure are taken from
\citet{Kauffmann17a} and \citet{Kauffmann17b} and references therein
(see Fig 4 of \citealt{Kauffmann17b}).
The gray dotted lines shown in the bottom panels correspond to the various annotated external pressures needed in order for
the gas to be in equilibrium, see Equation~\ref{eqn:v0}.
\label{fig:larsons_single}}
\end{figure*}

@@

% PVE
Motivated by observational studies, we also plot the $\sigma_{\rm gas}^2/R$ ratio and gas surface density of MCCs in the accreting phase of \flower in the bottom left panels of \Fig{larsons_single}. In the same plot, we show those observed in the Galactic Ring Survey (GRS) of the Milky Way \citep{Heyer09a} for comparison. Dashed lines in the figure show the loci along which the annotated external pressures are needed for any molecular clouds in equilibrium to have certain linewidths for a given set of surface densities (see \Sec{PVE}). 
At a given gas surface density, MCCs display a range of $\sigma^2/R$. 
Such variation in $\sigma^2/R$ is observed in GMCs in the central and outer part of the Milky Way \citep{Oka01a, Heyer09a}.

If constant column density (Larson's third law) is truly a fundamental properties of molecular clouds % or the roughly linear relationship between $\rho$ and $R$ of MCCs.
(i.e., if virialization is a universal property of molecular clouds with negligible surface pressure), 
one would expect a single point in the $\sigma_{\rm gas}^2/R$--$\Sigma$ relation at a given gas mass surface density $\Sigma$. 
The variation in $\sigma^2/R$ with $\Sigma$ seen here suggests that column density is not a fundamental properties of 
molecular clouds. Previous studies suggest that the observed mass-size relation of molecular clouds 
is solely a result of $\rho\propto M R^{-3}$ and may be an artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
%For instance, \citet{Ballesteros02a} show that the density
%of clouds ($\rho$) is independent of their size ($R$) when one properly extract these quantities in three-dimensional space instead of
%two-dimensional space; the latter is done in \obs.
%When they project the clouds into two-dimensional space and measure the densities and sizes of clouds for a given
%line tracer, they find $\rho \propto R^{-1}$, which is unsurprising since $N_H$ is just a constant.
%They therefore argue that the mass-size relation is simply a result of $\rho\propto M R^{-3}$
%observed in 2D-space at a fixed column density (cf. an actual physical relation of $M \propto R^{4/3}$ claimed in
%observational studies). % where $\rho$ would be proportional to $R^{-1/2}$
%% since $M$ would then scale as $R^{3}$ (ånd naturally display a linear relationship in $\log M - \log R$ plot).

@@

%
The gas velocity dispersion $\sigma_{\rm gas}$ of MCCs, on the other hand, is rather insensitive to the actual value of $n_{\rm cut}$.
%%
%In addition, for the densest MCC in the main disk of \flower, we find that while its size decreases as we increase $n_{\rm cut}$ --- as one would intuitively expect, the velocity dispersion remain approximately $\sigma\simeq$\,200\,\kms (see also \Fig{larsons_single}).
%%
This lack of variation is reassuring: our inference on the velocity dispersion of \z$\sim$\,6 MCCs in relation to those observed in
nearby and \z$\sim$2 galaxies is not biased by our choice of $n_{\rm cut}$ in this work.

@@

The variation in $\sigma^2/R$ with $\Sigma$ seen in this work is in line with previous studies, suggesting that ``Larson's third law''  
is an artifact arising from observational bias artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
This would imply column density is not a fundamental properties of molecular clouds and not all molecular clouds are virialized.

MCCs of \flower are found to have velocity dispersion and gas surface density systematically higher than
Milky Way clouds regardless of the density threshold $n_{\rm cut}$ adopted.
These MCCs are, in fact, highly supersonic, with high Mach number of $\bar{\mathcal{M}}\simeq6$.
Their velocity dispersions are comparable to those observed in $z$\ssim2 starburst galaxies ($\sigma_{\rm gas}\simeq200$\,\kms).
A comparison between the bulk and non-thermal velocity dispersions of MCCs indicates that most MCCs, excluding the arms and the disk of \flower, are supported by turbulence motions.

>>>>

\begin{abstract}
We study the properties of molecular cloud complexes (MCCs) at the Epoch of Reionization using cosmological zoom-in simulation.
We identify MCCs in a $z\simeq 6$ prototypical galaxy (``\flower'') 
using an H$_2$ density-based clump finder. 
We compare their mass, size, velocity dispersion, gas surface density, and virial parameter to \obs and perform Toomre-Q stability 
analysis to understand the origin and dynamical properties of MCC.
%
In \flower, the typical MCC mass is $10^{6.5}$\,\Msun, with sizes $R\simeq45-100$\,pc, 
which are comparable to nearby spirals and starburst galaxies. 
MCCs are highly supersonic and supported by non-thermal turbulence, with velocity dispersions of $\sigma_{\rm gas}\simeq$\,20\,--\,100\,\kms
and pressure of $\log (P/k_B)\simeq$\,7.6\,K\,\cc (i.e., $>$\,1000$\times$ the Milky Way).
That said, their velocity dispersion and gas surface density are comparable to nearby and $z$\ssim2 gas-rich starburst galaxies.
%
We identify gravitationally unstable MCCs via Toomre $Q_{\rm eff}$ analysis and
the more observationally accessible virial parameter ($\alpha_{\rm vir}$) analysis.
MCCs in the main disk of \flower are globally stable. MCCs in the densest regions have low $\alpha_{\rm vir}$ 
and $Q_{\rm eff}$ values, indicative of instability to collapse. 
Star formation is expected to take place within the clumps and cores on smaller scales, where non-thermal energy 
quickly dissipates and the gas structures are no longer supported by turbulence.
%
As such, spatial resolution of $\lesssim$40\,pc 
is necessary to shed light on the star-forming gas dynamics since the cosmic dark ages which is 
within reach with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA).
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

%mm rearranged and combined this and the following paragraph to proceed from observation to conclusion
Early galaxies have higher star formation rates \citep[SFR; ][]{Behroozi13b, Sparre15a, Maiolino15a, Dunlop17a} and smaller sizes \citep[e.g.,][]{Bouwens11a, Ono13a} compared to the local population \citep[see also a review by][]{Stark16a}.
%
As a consequence, we expect them to be significantly more ionized, and have intense and hard interstellar radiation fields (ISRF). Since their metallicity and dust content are also expected to be lower in these early evolutionary stages, shielding of UV photons---responsible for the photoheating of the gas---  is strongly reduced. Such differences in turn affect the regulation of the thermal and chemical state of the multi-phase interstellar medium (ISM). Studying ISM properties of early galaxies is essential for understanding how \SF proceeds under more extreme conditions.
%
Even in the local Universe, where detailed \obs can be made, variations in molecular cloud
%       \MM{do Hughes et al. study cloud complexes, as you do, or can you derive info about complexes from their data?}
%     \DL{This is just a statement to say there are variations among local GMCs. Together with the different galaxy properties found for early galaxies, it is interesting to study properties of molecular structures at earlier epochs. }
properties have been observed between different galaxy populations (see e.g., \citealt{Hughes10a, Hughes13b}).  Given that \highz galaxies statistically represent the early evolutionary stages of present-day galaxies, it is thus reasonable to pose the question: 
{\it what are the physical properties of
molecular cloud structures in early galaxies, and how do they differ from those found in local \galpop? }
%

@@

Motivated by Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, we assume equilibrium (i.e.,\ ${\ddot{\mathcal{I}}}=0$), define the cloud surface density as
$\Sigma$\eq$M/\pi R^2$, and rewrite the previous equation as
\begin{equation}
\frac{\sigma_{\rm gas}^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \frac{3}{5} \pi G \Sigma \right)\,,
\label{eqn:v0}
\end{equation}
which further reduces to
\begin{equation}
\frac{\sigma_{\rm gas}^2}{R} = \frac{\pi}{5} G \Sigma\,,
\label{eqn:SVE}
\end{equation}
if the external pressure $P_{\rm ext}=0$. For this case (often referred to as simple virial equilibrium) from the balance between kinetic and gravity terms we can define the virial parameter as
\begin{equation}
\alpha_{\rm vir} \equiv  \frac{5\sigma_{\rm gas}^2R}{GM_{\rm gas}} = \frac{5\sigma_{\rm gas}^2}{\pi G \Sigma R}\,,
\label{eqn:alpha}
\end{equation}

@@

In the clump-finding process (in position-position-position space; PPP), we employ a set of different density thresholds defined based on the molecular hydrogen density of \flower at different evolutionary stages ($z=6.0$--7.2).
%
We note that this process is the three-dimensional analog to identifying molecular structures based on the noise levels in position-position-velocity (PPV) maps that observers obtain with telescopes, using molecular line tracers such as CO, CS, and HCN. This is commonly done by
identifying clumps based on/after applying S/N-clipping, using tools such as \ncode{aips}'s task \ncode{serch}, \ncode{clumpfind}, and \ncode{cprops}; e.g., \citealt{Williams94a, Oka01a, Rosolowsky06a, Rosolowsky08a, DonovanMeyer13a}).
%
%We note that, owing to the nature of \obs, such structures are identified in position-position-velocity (PPV) space, whereas in simulations, one has the full 6D spatial-kinematic information, and can therefore cleanly identify structures directly using the density field in position-position-position (PPP) space.
Existing studies find a good correspondence in the dynamical properties extracted in PPV- versus PPP-space for well-isolated structures (\citealt{Ballesteros-Paredes02a, Heitsch09a, Shetty10a, Beaumont13a, Pan15a}, but see \citealt{Ballesteros-Paredes02a} and \citealt{Shetty10a} for a discussion on caveats and limitations).
% large scale structure in PPP may be identified as numerous lower mass structures in the PPV cube due to gradients in the LOS and conversly, superposition when spatially distinct components with the same velocity are "merged" into one component in the PPV space when observed along the LOS (see Fig. 1 of Beaumount13a). (may be worth putting more thoughts on this after the report is due).

@@

\begin{figure*}
\centering
\includegraphics[trim=0 0 0 0, clip, width=\textwidth]{\figpath/3by2_clumpProp_ss16-ss22.png}
\caption{
Linewidth-size relation (top), $\alpha_{\rm vir}$-mass relation
(middle), and $\sigma_{\rm gas}^2/R$-$\Sigma_{\rm gas}$ relation (bottom) for
MCCs (star symbols) identified in the two most extreme evolutionary
stages of \flower\ --- accreting phase (left) and starburst phase
(right). Star symbols are color-coded by the density thresholds
$n_{\rm cut}$, as illustrated by the colorbar.
Literature data in the linewidth-size plots are from \citet{Heyer04a, Rosolowsky05a, Bolatto08a,
% black dot = extragalactic GMCs in quiescent environments
Leroy15a}, and \citet{Swinbank11a}, and the empirical scaling relations are
from \citet{Larson81a, Heyer04a, Bolatto08a}.
Data points in the $\alpha_{\rm vir}$-mass figure are taken from
\citet{Kauffmann17a} and \citet{Kauffmann17b} and references therein
(see Fig 4 of \citealt{Kauffmann17b}).
The gray dotted lines shown in the bottom panels correspond to the various annotated external pressures needed in order for
the gas to be in equilibrium, see Equation~\ref{eqn:v0}.
\label{fig:larsons_single}}
\end{figure*}

@@

% PVE
Motivated by observational studies, we also plot the $\sigma_{\rm gas}^2/R$ ratio and gas surface density of MCCs in the accreting phase of \flower in the bottom left panels of \Fig{larsons_single}. In the same plot, we show those observed in the Galactic Ring Survey (GRS) of the Milky Way \citep{Heyer09a} for comparison. Dashed lines in the figure show the loci along which the annotated external pressures are needed for any molecular clouds in equilibrium to have certain linewidths for a given set of surface densities (see \Sec{PVE}). 
At a given gas surface density, MCCs display a range of $\sigma_{\rm gas}^2/R$. 
Such variation in $\sigma_{\rm gas}^2/R$ is observed in GMCs in the central and outer part of the Milky Way \citep{Oka01a, Heyer09a}.

If constant column density (Larson's third law) is truly a fundamental properties of molecular clouds % or the roughly linear relationship between $\rho$ and $R$ of MCCs.
(i.e., if virialization is a universal property of molecular clouds with negligible surface pressure), 
one would expect a single point in the $\sigma_{\rm gas}^2/R$--$\Sigma$ relation at a given gas mass surface density $\Sigma$. 
The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen here suggests that column density is not a fundamental properties of 
molecular clouds. Previous studies suggest that the observed mass-size relation of molecular clouds 
is solely a result of $\rho\propto M R^{-3}$ and may be an artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
%For instance, \citet{Ballesteros02a} show that the density
%of clouds ($\rho$) is independent of their size ($R$) when one properly extract these quantities in three-dimensional space instead of
%two-dimensional space; the latter is done in \obs.
%When they project the clouds into two-dimensional space and measure the densities and sizes of clouds for a given
%line tracer, they find $\rho \propto R^{-1}$, which is unsurprising since $N_H$ is just a constant.
%They therefore argue that the mass-size relation is simply a result of $\rho\propto M R^{-3}$
%observed in 2D-space at a fixed column density (cf. an actual physical relation of $M \propto R^{4/3}$ claimed in
%observational studies). % where $\rho$ would be proportional to $R^{-1/2}$
%% since $M$ would then scale as $R^{3}$ (ånd naturally display a linear relationship in $\log M - \log R$ plot).

@@

%
The gas velocity dispersion $\sigma_{\rm gas}$ of MCCs, on the other hand, is rather insensitive to the actual value of $n_{\rm cut}$.
This lack of variation is reassuring: our inference on the velocity dispersion of \z$\sim$\,6 MCCs in relation to those observed in
nearby and \z$\sim$2 galaxies is not biased by our choice of $n_{\rm cut}$ in this work.

@@

The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen in this work is in line with previous studies, suggesting that ``Larson's third law''  
is an artifact arising from observational bias artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}.
This would imply column density is not a fundamental properties of molecular clouds and not all molecular clouds are virialized.

MCCs of \flower are found to have velocity dispersion and gas surface density systematically higher than
Milky Way clouds regardless of the density threshold $n_{\rm cut}$ adopted.
These MCCs are, in fact, highly supersonic, with high Mach number of $\bar{\mathcal{M}}\simeq6$.
Their velocity dispersions are comparable to those observed in $z$\ssim2 starburst galaxies.
A comparison between the bulk and non-thermal velocity dispersions of MCCs indicates that most MCCs, excluding the arms and the disk of \flower, are supported by turbulence motions.
